---
title: "Plot `ANGSD` results"
author: 
  - "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  highlight: tango
  number_sections: no
  theme: default
  toc: yes
  toc_depth: 3
  toc_float:
    collapsed: no
    smooth_scroll: yes
---



# Setup

Setup R env. Load packages and set default image export formats, size and resolution.
```{r setup}
library(tibble)
library(ggplot2)
library(heatmaply)
library(ggdendro)
library(reshape2)
library(RColorBrewer)
library(cowplot)
options(scipen = 999) #Prevent scientific notation
```

File names and important variables.
```{r file_names}
annot.file <- "Annotations.tsv"
admixture.file <- "PCAngsd.Admixture.tsv"
allelFreqs.file <- "PCAngsd.IndAlleleFreq.tsv"
```



# Load data 

*Annotations*
```{r load_metadata}
annot <- read.table(annot.file, header=T, sep='\t', comment.char='') %>% 
  column_to_rownames(var = "sample_id")
```

*Color palette*
```{r color_palette}
annot.labels <- annot[c("group_id","best_ploidy_model")]
annot.cols <- annot[c("group_id_color","best_ploidy_model_color")] %>% 
  rename("group_id" = "group_id_color") %>% 
  rename("best_ploidy_model" = "best_ploidy_model_color")

# Get color combinations for plotting
palette <- as.character(c(
  annot.cols$group_id, 
  annot.cols$best_ploidy_model
))
names(palette) <- as.character(c(
  annot.labels$group_id, 
  annot.labels$best_ploidy_model
))
palette <- palette[!duplicated(palette)]
```

*Admixture*
```{r load_Admixture_data}
admixture <- read.table(admixture.file, header=T, sep='\t', comment.char='', check.names=FALSE) %>% 
  column_to_rownames(var = "sample_id")
```

*IndAlleleFreq*
```{r load_IndAlleleFreq_data}
allelFreqs <- read.table(allelFreqs.file, header=T, sep='\t', comment.char='', check.names=FALSE) %>% 
  column_to_rownames(var = "sample_id")
```



# Plot Admixture based on two PC

A function that wraps up all the functions required to parse and plot the input data.
Will use the supplied hclust object to create dendrogram and order bars or will perform clustering itself.
``` {r func_cluster_plot}
cluster_plot <- function(data, samples, selected_columns, 
                         plot.dendro=NULL, colors=NULL,
                         dist.method="euclidean", hclust.method="complete",
                         x.label.font.size=5, legend.size.scale=0.5,
                         legend.title="Groups",
                         rel_heights=c(0.3, 1.0, 0.3)) {
  # Code modified from: https://stackoverflow.com/questions/44646488/stacked-bar-plot-with-hierarchical-clustering-dendrogram
  
  #########################
  #### Cluster samples ####
  #########################
  if(is.null(plot.dendro)) {
    d <- dist(as.matrix(data), method=dist.method)
    hc <- hclust(d, method=hclust.method)
    plot.dendro <- as.dendrogram(hc)
  }
  
  ##############
  #### Plot ####
  ##############
  # Convert dendrogram to segment data
  dend_data <- dendro_data(plot.dendro, type="rectangle")
  segment_data <- dend_data[["segments"]]
  dend_labels <- gsub('\'', '', dend_data$labels$label)

  # Sample positions df
  sample_pos_table <- with(dend_data$labels, 
                           data.frame(x_center = x, sample = as.character(label), width = 0.9))
  # X-axis limits
  axis_limits <- with(sample_pos_table, 
                      c(min(x_center - 0.5 * width), max(x_center + 0.5 * width))) + 
                      0.1 * c(-1, 1) # extra spacing: 0.1
  
  # Plot clustering dendrogram
  p.dendr <- ggplot(segment_data) +
    geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
    scale_y_continuous(expand = c(0, 0, 0, 0.1)) + 
    scale_x_continuous(breaks = sample_pos_table$x_center, 
                       labels = NULL, 
                       limits = axis_limits, 
                       expand = c(0, 0)) + 
    labs(x = "", y = "Distance", colour = "", size = "") +
    theme_bw() + 
    theme(panel.grid.minor=element_blank(),
          panel.grid.major=element_blank(),
          plot.margin=unit(c(0,0,0,0), "mm"))
  
  # Plot bar chart
  p.barplot <- data %>%
    rownames_to_column() %>%
    arrange(as.character(factor(rowname, levels = dend_labels)), dend_labels) %>%
    column_to_rownames("rowname") %>%
    as.matrix() %>%
    melt() %>%
    rename(Sample=Var1, ITS2type=Var2, Abundance=value) %>%
    # Plot bar chart
    ggplot(aes(x=Sample, y=Abundance, fill=ITS2type)) +
      geom_bar(position="stack", stat="identity") +
      scale_y_continuous(expand = c(0, 0)) +
      scale_fill_manual(values = colorRampPalette(brewer.pal(12,"Paired"))(12)) +
      labs(x = "", y = "Frequency", fill=legend.title) +
      theme_bw() +
      theme(axis.text.x=element_blank(),
            legend.key.height= unit(2*legend.size.scale, 'mm'),
            legend.key.width= unit(2*legend.size.scale, 'mm'),
            legend.title = element_text(size=4*legend.size.scale),
            legend.text = element_text(size=4*legend.size.scale),
            plot.margin=unit(c(0,0,0,0), "mm")) +
      guides(fill=guide_legend(ncol=1))
  
  # Plot Meta Information about samples
  if(is.null(colors)) {
    colors <- rep(brewer.pal(brewer.pal.info["Set3", "maxcolors"], "Set3"), ceiling(nrow(samples)/12))
  }
  p.sampleInfo <- samples %>%
    select(all_of(selected_columns)) %>%
    rownames_to_column() %>% 
    arrange(as.character(factor(rowname, levels = dend_labels)), dend_labels) %>% 
    column_to_rownames("rowname") %>% 
    as.matrix() %>% 
    melt() %>% 
    rename(Sample=Var1, MetaInfo=Var2, Value=value) %>% 
    mutate(Count = 1) %>% 
    filter(MetaInfo!="PlugID") %>%
    ggplot(aes(x=Sample, y=Count, fill=Value)) + 
      geom_bar(position="stack", stat="identity") + 
      scale_y_continuous(expand = c(0, 0)) + 
      scale_fill_manual(values=colors) +
      labs(x = "Samples", y = "Meta Information", fill="") +
      theme_bw() +
      theme(axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1, size=x.label.font.size),
            axis.ticks.y=element_blank(),
            axis.text.y=element_blank(),
            legend.key.height= unit(2*legend.size.scale, 'mm'),
            legend.key.width= unit(2*legend.size.scale, 'mm'),
            legend.title = element_text(size=4*legend.size.scale),
            legend.text = element_text(size=4*legend.size.scale),
            plot.margin=unit(c(0,0,0,0), "mm")) +
      guides(fill=guide_legend(ncol=1))
  
  # Combine plots and print to screen
  comb <- plot_grid(p.dendr, p.barplot, p.sampleInfo, align = 'v', ncol = 1, axis = "lr", rel_heights = rel_heights)
  print(comb)
}
```



Plot Admixture results. Use the admixture matrix to generate the degdrogram and sample order.
```{r plot_Admixture}
cluster_plot(admixture, annot.labels, c("group_id", "best_ploidy_model"), colors=palette, x.label.font.size=2)
```




# PCA using PCAngsd results with estimated individual allele frequencies

Generate eigen vectors.
```{r load_PCAngsd_IndAllelFreq_data}
C <- as.matrix(allelFreqs)
e <- eigen(C)
df <- as.data.frame(e$vectors)
rownames(df) <- rownames(C)
df <- cbind(df, annot.labels)
```


Plot PC1 and PC2 of PCAngsd results WITH estimated individual allele frequencies using the `plotly` package. 
```{r plotly_PCAngsd_IndAllelFreq_PC1_PC2}
# group_id
p <- plot_ly(df, x=~V1, y=~V2, text=rownames(df), type = "scatter",
             mode="markers", color=~group_id, colors=palette, marker=list(size=11))
p <- layout(p,title="Without individual allele frequency - colored by group_id",
            xaxis=list(title="PC1"),
            yaxis=list(title="PC2"))
p

# Reef
p <- plot_ly(df, x=~V1, y=~V2, text=rownames(df), type = "scatter",
             mode="markers", color=~best_ploidy_model, colors=palette, marker=list(size=11))
p <- layout(p,title="Without individual allele frequency - colored by best_ploidy_model",
            xaxis=list(title="PC1"),
            yaxis=list(title="PC2"))
p
```


# Session Info

```{r ressionInfo}
sessionInfo()
```

